---
title: "P8105 Homework 3"
output: github_document
date: "October 15th, 2022"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(lubridate)
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.dim = c(12, 7))
```

## Problem 1: Instacart Data

We will be working with the “The Instacart Online Grocery Shopping Dataset 2017”. 

### Data overview and description 
```{r dataload}
data("instacart")

instacart
```
The dataset contains information about online grocery orders made through Instacart. There 1,384,617 observations of 15 variables, where each row in the dataset is a product from an order. Each order is associated with a unique order ID number and order ID. The variables `order_hour_of_day` and `order_dow` describe the time (hour of day and day of the week) orders were made. Each row also contains information about each product, such as `product_name`, `aisle`, and `department` of the product. The variable `reordered` indicates whether a product has been purchased by a user in the past, and `add_to_cart_order` indicates the order an item was placed in the cart.

For example, for `order_id == 1`, we can see the order was made at 10AM on a Thursday, 4 out of the 8 products are reorders, and that most of the products came from the produce and dairy eggs department.

## Data exploration

Now, we want to conduct some basic EDA and answer some questions about the `instacart` data.

1. Number of aisles

Using `group_by`, and `summarise`, we can determine the number of aisles, and the most popular aisles customers order from in the dataset.

```{r}
aisles = instacart %>% 
  group_by(aisle_id, aisle) %>% 
  summarise(
    n_obs = n()
  ) %>% 
  arrange(desc(n_obs))

aisles
```

There are `r nrow(aisles)` different aisles. The most popular aisles customers order from are fresh vegetables, fresh fruits, and packaged vegetables fruits, respectively.

2. Plotting number of items ordered in each aisle

```{r plot_aisles}
plot_aisles = aisles %>% 
  filter(n_obs >= 10000) %>% 
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(
    title = "Number of Instacart items ordered by aisle",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Data from The Instacart Online Grocery Shopping Dataset 2017 in the p8105.datasets package"
  )

plot_aisles
```

Through the bar plot, we can confirm that for aisles with over 10,000 items ordered, the most popular aisles are fresh fruits and vegetables, and the least popular aisles are butter, oils and vinegars, and dry pasta.

3. Popular items 

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r pop_items}
pop_items = instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(n_obs = n()) %>% 
  arrange(desc(n_obs),.by_group = TRUE) %>% 
  slice(1)

knitr::kable(pop_items)
```


4. Mean ordering times

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

## Problem 2: Accelerometer Data

### Clean and load data

In this problem, we will load, clean, and tidy the `accel_data.csv` dataset, which contains five weeks of accelerometer data collected on a 63 year-old male with BMI 25, admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). 

We will apply snake case to all the variables, add a logical `weekend` variable that is `TRUE` if `day == c("Saturday", "Sunday")` and `FALSE` otherwise. 

```{r accel_data}
accel_data =
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    weekend = ifelse(day == "Saturday" | day == "Sunday", TRUE, FALSE)) %>% 
  relocate(weekend, .after = day)

accel_data
```

The dataset contains 35 observations of 1444 variables, where each row describes the activity counts for all minutes of each day. The weekday of observation is indicated by `day`, and has a unique `day_id`. `week` indicates the week of observation, and the `weekend` indicates whether observation falls on a weekend. The variables `activity_*` are the activity counts for all 1440 minutes of a 24-hour day, starting at midnight.

### Aggregate across minutes

Using `rowSums` and `across` within the `mutate` function, we can compute the sum across minutes for each day to get the total activity over the day. 

```{r aggregate}
accel_data = accel_data %>%
  mutate(total_activity = rowSums(across(activity_1:activity_1440))) %>% 
  relocate(total_activity, .after = weekend) 

accel_data %>% 
  select(week:total_activity) %>% 
  knitr::kable()
```

The table above shows the totals for each day. From first glance, we see that the total activity is higher on weekends compared to the weekdays We can also see that as time progresses, activity levels appear to slightly decrease.

### Plotting activity through the day

Accelerometer data allows the inspection activity over the course of the day. Using `summarise`, we can get the mean activity counts for each minute of the day, for each day of the week. 

```{r daily_data}
daily_activity = accel_data %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity",
    names_prefix = "activity_") %>% 
  mutate(minute = as.numeric(minute)) %>% 
  group_by(day, minute) %>% 
  summarise(mean_activity = mean(activity))

```

Based on this data, we can plot the mean 24-hour activity time course, according to the day of the week.

```{r activity_plot}
activity_plot = daily_activity %>% 
  ggplot(aes(x = minute, y = mean_activity, color = day)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth(se = FALSE) + 
  labs(
    x = "Minute",
    y = "Mean activity",
    title = "Mean daily accelerometer activity by day of the week",
    caption = "Data come from the accel_data.csv dataset",
    color = "Day"
  ) + 
  scale_x_continuous(limits = c(0, 1400)) + 
  scale_y_continuous(limits = c(0, 2300)) + 
  viridis::scale_color_viridis(discrete = TRUE)

activity_plot
```

Describe in words any patterns or conclusions you can make based on this graph.