---
title: "P8105 Homework 3"
output: github_document
date: "October 15th, 2022"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Problem 1: Instacart Data

We will be working with the “The Instacart Online Grocery Shopping Dataset 2017”. 

### Data overview and description 
```{r dataclean}
data("instacart")

instacart
```
The dataset contains information about online grocery orders made through Instacart. There 1,384,617 observations of 15 variables, where each row in the dataset is a product from an order. Each order is associated with a unique order ID number and order ID. The variables `order_hour_of_day` and `order_dow` describe the time (hour of day and day of the week) orders were made. Each row also contains information about each product, such as `product_name`, `aisle`, and `department` of the product. The variable `reordered` indicates whether a product has been purchased by a user in the past, and `add_to_cart_order` indicates the order an item was placed in the cart.

For example, for `order_id == 1`, we can see the order was made at 10AM on a Thursday, 4/8 products are reorders, and that most of the products came from the produce and dairy eggs department.


